----
title: "Preferential Attachment Analysis"
output: html_document
---

## Packages

```{r message=FALSE, warning=FALSE}
library(compiler)
library(dplyr)
library(tidygraph)
library(ggraph)
library(tidyverse)
library(sna)
library(igraph)
library(visNetwork)
library(networkD3)
library(zeallot)
```

## Import network data

The dataset has two dataframes that we can use to get the information about nodes and the information about edges. We rename the columns of the matchup dataframe to fit the edges definition and we have an id to the lineup dataframe to refer to lineup players by their node id.

The original dataset of matchups has lineups numbered from zero, but tidygraph (and most packages in R) have indices that start in 1. For that reason we add 1 to the entries of the edges table and also to lineup table.

```{r message=FALSE, warning=FALSE}
lineup <- read_csv("lineups.csv")
lineup$id <- lineup$id +1 
names(lineup) <- c('id','label','team')

matchups <- read_csv("matchups.csv")
matchup <- matchups[,c("home_players","away_players","type","weight","home_team","away_team","time")]
names(matchup) <- c('from','to','weight','score',"home_team","away_team","time")
matchup$to <- matchup$to+1
matchup$from <- matchup$from+1
matchup <- matchup[!(matchup$weight==0),]


#consider time effect
matchup$score_adjusted <- matchup$score/matchup$time 

matchup$weight <- factor(matchup$weight)

matchup$RowID <- seq.int(nrow(matchup))


print(paste("# Lineups: ",nrow(lineup)))
print(paste("# Matchups: ",nrow(matchups)))
print(paste("# Matchups after filtering: ",nrow(matchup),"(differences:",nrow(matchups)-nrow(matchup),")"))
```



## Preferential Attachment for Home/away team lineups

### Home team lineups analysis

Order the matchup table by the id of Home team lineup and treat the time steps as the times that a lineup showed up as a home team. For example, if lineup A showed up as Home team at the first time step, we treat this matchup as if it happened at t_0. Then the second time corresponds to t_1.  

```{r message=FALSE, warning=FALSE}
# pre-analysis
home_matchup <- matchup[order(matchup$from),] %>% 
  group_by(from) %>% 
  mutate(time_step = 1:n()) %>% ungroup()
head(home_matchup)

# Histogram of counts
hist(count(home_matchup,from)[count(home_matchup,from)$n>1,]$n,xlim=c(0,50),freq = T,main="Histgram of counts",xlab="counts",breaks=150)
count(home_matchup,from) %>%  ggplot(aes(x = from, y = n)) + geom_col()
```

Add positive outdegree at each time_step for each home lineups. Save results to an RData file.
```{r message=FALSE, warning=FALSE}
home_counts <- count(home_matchup,from)
home_matchup_degree <- home_matchup[0,] %>% add_column(deg_pout = NA)

for (i in c(1:nrow(home_counts))) {
  id <- home_counts$from[[i]]
  temp <-  home_matchup[home_matchup$from == id, ] %>% add_column(deg_pout = NA)
  temp$weight <- as.numeric(temp$weight)
  temp$weight[temp$weight == 1] <- 0
  temp$weight[temp$weight == 2] <- 1
  for (j in c(1:nrow(temp))){
    temp[j,"deg_pout"] <- sum(temp$weight[1:j])
  }
  home_matchup_degree <- rbind(home_matchup_degree,temp)
}

save(home_matchup_degree, file = "RData/home_matchup_degree.RData")
```

Calculate the probability of a home lineup win at the current time step if they have positive outdegree = past_deg. 
```{r message=FALSE, warning=FALSE}
load("RData/home_matchup_degree.RData")
attraction_home <- function(past, now, past_deg,now_deg){
    data <- home_matchup_degree %>% 
      inner_join(home_matchup_degree %>% filter(time_step==past) %>%select(from)) %>%
      inner_join(home_matchup_degree %>% filter(time_step==now) %>% select(from)) %>%
      filter(time_step == past | time_step == now)
    
    data_denominator <- data %>% filter(deg_pout ==past_deg & time_step==past)
    data_numerator <- data %>% filter(deg_pout== now_deg & time_step==now) %>%
        inner_join(data_denominator %>% select(from))

    return(list(nrow(data_numerator),nrow(data_denominator)))
}
```


Combine the results in a dataframe.
Takes ~ 2 hours.
```{r message=FALSE, warning=FALSE}
home_counts <- count(home_matchup_degree,from) # count the number of time steps for each node
home_win_prob <- data.frame(past_deg=as.numeric(), now_deg = as.numeric(), win_prob = as.numeric(), n1 =as.numeric(), n2 = as.numeric(), stringsAsFactors=FALSE)


for (i in c(1:(max(home_matchup_degree$deg_pout)-1))){
  if (i %% 10 == 0){
    print(i)
  }
  home_win_prob[i, "past_deg"] <- i
  home_win_prob[i, "now_deg"] <- i+1
  n1 <- 0
  n2 <- 0
  deg1 <- home_win_prob$past_deg[[i]]
  deg2 <- home_win_prob$now_deg[[i]]
  for (j in c(2:max(home_counts$n))){
      c(a,b) %<-% attraction_home(past = j-1, now = j, past_deg = i,now_deg = i+1)
      n1 <- n1 + a
      n2 <- n2 + b
    }
  home_win_prob[i,"win_prob"] <- n1/n2
  home_win_prob[i,"n1"] <- n1
  home_win_prob[i,"n2"] <- n2
}

save(home_win_prob, file = "RData/home_win_prob.RData")
```


```{r message=FALSE, warning=FALSE}
load("RData/home_win_prob.RData")


reg_home <- lm(win_prob ~ past_deg, data = home_win_prob[home_win_prob$n2>70,])
summary(reg_home)

plot(x=home_win_prob$past_deg,y=home_win_prob$win_prob,main = "Probability of winning with home lineups",xlab = "Past wins", ylab = "Odds of winning")+abline(v=home_win_prob$past_deg[home_win_prob$n2<70][1],col="blue")+abline(reg_home,col="red",lwd=3, lty=2)
```



### Away team lineups analysis

Order the matchup table by the id of Away team lineup and treat the time steps as the times that a lineup showed up as an away team. For example, if lineup A showed up as away team at the first time step, we treat this matchup as if it happened at t_0. Then the second time corresponds to t_1.  


```{r message=FALSE, warning=FALSE}
# pre-analysis
away_matchup <- matchup[order(matchup$to),] %>% 
  group_by(to) %>% 
  mutate(time_step = 1:n()) %>% ungroup()
head(away_matchup)

count(away_matchup,to) %>%  ggplot(aes(x = to, y = n)) + geom_col()
```


Add the positive indegree at each time_step for each away lineup. Save results to an RData file.
```{r message=FALSE, warning=FALSE}
away_counts <- count(away_matchup,to)
away_matchup_degree <- away_matchup[0,] %>% add_column(deg_pin = NA)

for (i in c(1:nrow(away_counts))) {
  id <- away_counts$to[[i]]
  temp <-  away_matchup[away_matchup$to == id, ] %>% add_column(deg_pin = NA)
  temp$weight <- as.numeric(temp$weight)
  temp$weight[temp$weight == 1] <- 1
  temp$weight[temp$weight == 2] <- 0
  for (j in c(1:nrow(temp))){
    temp[j,"deg_pin"] <- sum(temp$weight[1:j])
  }
  away_matchup_degree <- rbind(away_matchup_degree,temp)
}

save(away_matchup_degree, file = "RData/away_matchup_degree.RData")
```

Calculate the probability of an away lineup win at current time step if they have a positive indegree = past_deg. 
```{r message=FALSE, warning=FALSE}
load("RData/away_matchup_degree.RData")
attraction_away <- function(past, now, past_deg,now_deg){
    data <- away_matchup_degree %>% 
      inner_join(away_matchup_degree %>% filter(time_step==past) %>%select(to)) %>%
      inner_join(away_matchup_degree %>% filter(time_step==now) %>% select(to)) %>%
      filter(time_step == past | time_step == now)
    
    data_denominator <- data %>% filter(deg_pin ==past_deg & time_step==past)
    data_numerator <- data %>% filter(deg_pin== now_deg & time_step==now) %>%
        inner_join(data_denominator %>% select(to))

    return(list(nrow(data_numerator),nrow(data_denominator)))
}
```


Combine the results in a dataframe.
Takes ~ 2 hours.
```{r message=FALSE, warning=FALSE}
away_counts <- count(away_matchup_degree,to) # count number of time steps for each node
away_win_prob <- data.frame(past_deg=as.numeric(), now_deg = as.numeric(), win_prob = as.numeric(),n1 =as.numeric(), n2 = as.numeric(),  stringsAsFactors=FALSE)


for (i in c(1:(max(away_matchup_degree$deg_pin)-1))){
  away_win_prob[i, "past_deg"] <- i
  away_win_prob[i, "now_deg"] <- i+1
  n1 <- 0
  n2 <- 0
  deg1 <- away_win_prob$past_deg[[i]]
  deg2 <- away_win_prob$now_deg[[i]]
  for (j in c(2:max(away_counts$n))){
      c(a,b) %<-% attraction_away(past = j-1, now = j, past_deg = i,now_deg = i+1)
      n1 <- n1 + a
      n2 <- n2 + b
    }
  away_win_prob[i,"win_prob"] <- n1/n2
  away_win_prob[i,"n1"] <- n1
  away_win_prob[i,"n2"] <- n2
}

save(away_win_prob, file = "RData/away_win_prob.RData")
```


Regression analysis
```{r message=FALSE, warning=FALSE}
load("RData/away_win_prob.RData")

reg_away <- lm(win_prob ~ past_deg, data = away_win_prob[away_win_prob$n2>70,])
summary(reg_away)

plot(x=away_win_prob$past_deg,y=away_win_prob$win_prob,main = "Probability of winning with away lineups",xlab = "Past wins", ylab = "Odds of winning")+abline(v=away_win_prob$past_deg[away_win_prob$n2<70][1],col="blue")+abline(reg_away,col="red",lwd=3, lty=2)
```

